### 2017-06-10 report/paper.tex
report update

\end{titlepage}
\tableofcontents
\newpage
\begin{abstract}
Common music recommender systems do not take into account the position of the customer. This piece discusses how it is possible to link a song with a POI using the semantic web.
\end{abstract}

\section{Introduction}
Linkng POIs and songs is a big challenge, in fact they are part of two completely separated concepts that cannot be easily linked also by a person it would be difficult to state some criteria that can link a song with a place. We present an approach that attempts to leverage this problem by using the semantic relations exposed by DBpedia.
\subsection{Motivation and context}

\subsection{Research problems}

@@

\subsection{Report structure}
The report is divided into four sections, one for each part of the project. The first two parts are about linking external datasets to Dbpedia, this is also known as enrichment and it consists of linking in the correct way two entities of different datasets
\section{Interlinking and enriching POIs}

\subsection{Problem description}
Finding relations inside DBpedia is a known problem which has been resolved in the past. So we could simply look for POIs and Artists in Dbpedia and then link them using known algorithms.
Unfortunately DBpedia is not perfect: it is true that it contains POIs and Artists, but these entities are not cleanly defined, they usually contains dirty data that can tremendously affect our performances.
POIs and Artists can be found in highly specialized datasets, but unfortunately they do not contain links to eachother, and even worse they are not linked to DBpedia. This means that by using these databases alone we cannot find relations.
What is done in this cases is to link highly specialized datasets to DBpedia, in this way we can exploit the advantages of both worlds: highly specialized datasets give us confidence and trust in our main entities (POIs and Artists), while DBpedia allow us to find relations between them thanks to its generality.
The following two sections will explain in a detailed way the whole process of connecting these highly specialized datasets to DBpedia.

@@

\subsection{Evaluation (Google Map / OpenStreetMap)}
The evaluation part is done using both the Google Map and OpenStreetMap. We calculated the Medium Square Error of the POIS to the positions recovered with the APIs, the error distance is the haversine distance.
\section{Interlinking and enriching Artists}
\subsection{Problem description}
\subsection{Datasets: DOREMUS, DBpedia, Wikidata}
\subsection{Entity Linking process (+ ElasticSearch)}
\subsubsection{Similarity measures}
\subsubsection{Aggregations}
\subsubsection{Filtering}
\subsection{Evaluation}
\section{Path finder}
\subsection{Problem description}
\subsection{Related work (relfinder)}
\subsection{Proposed approach and best path selection algorithm}

>>>>

\end{titlepage}
\tableofcontents
\newpage
\begin{abstract}
Abstract: In this project we show a new approach for finding artists linked to a place of interest (POI). We propose a semantic-based framework built upon the DBpedia, 3cixty, and Doremus. Furthermore we present a graph exploration algorithm that is able to find deep connections in an short time and discriminate between them by looking at those that can be the most interesting. Our project aims to making the user know about the existing connections that can link music to a Place, by stimulating the curiosity of the user proposing non common relations.
\end{abstract}

\section{Introduction}
todo: change
Most of the available music recommenders suggest music without taking into consideration the user’s context, e.g., her mood, current location, activity. 
Finding music that suits a POI can be viewed as a context-aware recommendation problem, the place being the context for consuming the recommendation.         
The main challenge that one must face when addressing the above mentioned goal is related to the fact that POIs and music are two rather different domains, and there is no obvious way to match such heterogeneous items. However, with the advent of the Semantic Web, and specifically with its reference implementation in the Linked Data initiative, new opportunities arise to face the above difficulties.

\subsection{Motivation and context}
In the past, many tried to link POIs and music in many different way. xxx managed to link music and POIs by using a common set of tags that describe emotional properties both for music and POIs.
This method is completely supervised, in fact it is necessary to define in a clean way emotions for songs and POIs. This operation was done by hand and is hardly scalable.
Marius Kaminskas restrict the subspace of DBpedia by identifying classes belonging to the two domains of interest, and the relations existing between instances of such classes. In this way they managed to find relations in a set of constrained paths. Looking on the future works proposition, they proposed the idea to exploit arbitrary semantic relations between POIs and musicians, e.g. direct relations such as 'Gustav Mahler was the director of Vienna State Opera', and complex non-directed relations such as 'Ana Bel' (a famous Spanish singer) composed a song whose lyrics are about La Puerta de Alcala (a well-known POI in Madrid, Spain). 
In order to do this, they wanted to use Relfinder and set the connections' weights programatically. In fact one of the bad things was that connections weights were assigned by experts and were not provided in the paper.

What we intend to do is to find relations in a completely unsupervised way: this means that we intend to define a pathfinder that finds paths and a path discriminator that selects the interesting algorithms. Our intention is to limit to zero the need of human intervention in the hope that this approach will be completely data driven and easily scalable

\subsection{Research problems}
Finding relations inside DBpedia is a known problem which has been solved in the past. So we could simply look for POIs and Artists in Dbpedia and then link them using known algorithms.
Unfortunately DBpedia is not perfect: it is true that it contains POIs and Artists, but these entities are not cleanly defined, they usually contains dirty data that can tremendously affect our performances.
POIs and Artists can be found in highly specialized datasets, but unfortunately they do not contain links to eachother, and even worse they are not linked to DBpedia. This means that by using these databases alone we cannot find relations.
What is done in this cases is to link highly specialized datasets to DBpedia, in this way we can exploit the advantages of both worlds: highly specialized datasets give us confidence and trust in our main entities (POIs and Artists), while DBpedia allow us to find relations between them thanks to its generality.
The following two sections will explain in a detailed way the whole process of connecting these highly specialized datasets to DBpedia.

@@

\subsection{Report structure}
The report is divided into five sections, one for each part of the project. The first two parts are about linking external datasets to Dbpedia, this is also known as entity linking and we perform it in the optic of enriching our data with the inter-concept links present in DBpedia.
The third treats the pathfinder, a module of our system that is able to find deep connections between two entities in DBpedia.
The fourth describes the whole implementation of the REST API that allows the application to work smoothly.
The fifth part is about the mobile web application and the technologies involved in its functioning.

\section{Interlinking and enriching POIs}
\subsection{Problem description}
Finding relations inside DBpedia is a known problem which has been solved in the past. So we could simply look for POIs and Artists in Dbpedia and then link them using known algorithms.
Unfortunately DBpedia is not perfect: it is true that it contains POIs and Artists, but these entities are not cleanly defined, they usually contains dirty data that can tremendously affect our performances.
We start our search of connecting POIs and Artists from the POIs. POIs can be found in highly specialized datasets that contains clean data that we can trust, one of these datasets is 3cixty. Now, it would be optimal to find a direct mapping between a 3cixty POI and a DBpedia POI, in this way we could actually enrich the 3cixty POI with the context informations that can be found on DBpedia. Unfortunately this link does not exist and we must create it on our own by using an heuristic.

@@

\subsection{Evaluation (Google Map / OpenStreetMap)}
The evaluation part is done using both the Google Map and OpenStreetMap. We calculated the Medium Square Error of the POIS to the positions recovered with the APIs, the error distance is the haversine distance.
\section{Interlinking and enriching Artists}
\subsection{Problem description}
\subsection{Datasets: DOREMUS, DBpedia, Wikidata}
\subsubsection{DOREMUS}
3cixty is a semantic dataset that contains aggregated informations of POIs and events. The dataset is the result of an aggregation process that join the POI informations from different datasets e.g. Facebook, Foursquare, Yelp \dots
In this project 3cixty is used as a trusted source of POIs for Nice.
POIs have many informations, such as the category, the address, comments from users, and coordinates. In order to get all the POIs in Nice, we simply used the POIs endpoint selecting the Nice city.
\subsubsection{DBpedia}
DBpedia is a project that aims at connecting the semantic world. We used it to get the POIs in Nice.
In order to get the POIs from DBpedia, it is necessary to perform SPARQL queries against the SPARQL endpoint. POIs are retrieved using two different sparql queries that target different kind of resources:
\begin{enumerate}
\item We perform a query that looks for all Geoentities that are in the area of nice. An entity is in the area of Nice if its coordinates are in a custom box that contains all the area of Nice.
Unfortunately there are many POIs that are without coordinates,so they are not selected with this query.
\item We perform some queries that get in a recursive way all entities that have as a parent the category of nice or nice and that are related to a place. The relation to a place is a list of classes that is done by hand and could be subjected to refinements.
\end{enumerate}
\subsubsection{Wikidata}
Wikidata is a project that aims at connecting the semantic world. We used it to get the POIs in Nice. Respect to DBpedia, Wikidata has a well defined ontology that is strictly enforced. This means that it is easier to select what we want. Moreover, most of POIs in wikidata have coordinates, so we can get them with a single query. Since wikidata supports complex geo-based queries, we made a query that looked for all POIs in a range of 25 Kilometers from the center of Nice.

\subsection{Entity Linking process (+ ElasticSearch)}
\subsubsection{Properties}
\subsubsection{Transformations}
\subsubsection{Similarity measures}
\subsubsection{Aggregations}
\subsubsection{Filtering}
\subsection{Evaluation}
todo: add source of isni
The evaluation of the found links is done by comparing the found links with those provided by ISNI. ISNI is the ISO certified global standard number for identifying the millions of contributors to creative works. By achieving these goals the ISNI will act as a bridge identifier across multiple domains and become a critical component in Linked Data and Semantic Web applications.

ISNI matches are made by humans, so we can consider them as a ground truth that can be trusted. By comparing our matches we want to see if our performances are acceptable and if we can discover new links between ISNI numbers and DBpedia.

todo: table with cool statistics
todo: graph image

We can state that the proposed method based on combinations of labels, classes and dates is able to match almost perfectly the artists of the two datasets. Moreover, we are able to find 2000 artists that at the time of writing are not present in the ISNI register.

\section{Path finder}
\subsection{Problem description}
todo: add citations
Once we have a direct mapping between entities that comes from specialized datasets and DBpedia, we actually need to link POIs and Artists.
The problem of finding relations between entities in DBpedia has already been approached and solved in the past, especially from Relfinder. The whole problem can be solved by writing a single query that finds a path of a given depth between two resources. This approach works well when datasets are relatively small and the two entities are two concepts that are semantically close. Instead, in our case we have concepts that are definetly far from each other, which implicates an high depth of an eventual path, and most importantly, we need to match hundreds of POIs with thousands of artists, This makes means that at the end the whole process will be repeated 100.000 times.
\subsection{Related work (Relfinder)}
The path finding problem has already been solved in the past with Relfinder: a tool that is able to find all paths with a specified depth between two entities. 
In fact, when we think at linking two entities, we need to set the position of the middle node and the number of middle nodes.

$$a-->b-->c-->d-->e=m$$
$$a-->b-->c-->d=m<--e$$
$$a-->b<--c=m-->d-->e$$

These are all different paths that links two entities 'a' and 'e'. In the simpler case, the arrows goes in one direction. In the second case they change direction one time.In the third case they change two times.
Relfinder manages to solve the problem by making different query that cover all possible combinations of the path specified above. As a constraint, in Relfinder the direction cannot change more than two times. There are still a huge amount of queries that explodes when the depth starts to be deep.
Here we intend as a deep depth a depth greater than 4, in fact this is like a magic threshold. 
Moreover, when we use Relfinder and set a greater depth we incur in two problems: 
\item The number of queries explodes because we need to consider all possible combinations
\item The time required by a single query explodes too, most of the times we cannot even do a query between two entities in a short time.
In order to succeed we need to solve those two problem and we need to create a fast implementation, because at the end we will need to do 100.000 queries. 
\subsection{Proposed approach}
The problem is that DBpedia is small world and its average degree is 4.3, this means that on average we can reach every entity in at most 4-5 hops. Unfortunately our problem does not make part of that average, and many times we need 5-6 hops to connect two entities. The problem is that when we make a query that looks for paths with depth 6 we are basically querying the whole DBpedia, which at the time of writing is composed of 8.8 billion triples.
It is our belief that a query that tries to find a path between two entities, is converted by the SPARQL engine in what we can compare to a join for each hop. This, with the small world characteristic of DBpedia means that we are doing huge joins for each query.
Another way to visualize the query is a Breath First Search that starts from the source entity, and it expands hop after hop until the specified depth is reached. In this way is even more easy to see why we and up by querying billions of triples.
A common improvement of BFS is the two-way BFS, where the same algorithm is applied both at the source and the destination. In this way we can see that we can reach the same middle node by querying a way smaller portion of the graph.
Then, looking by hand at paths, we have seen that usually one direction inversion cover most of the cases, by limiting this we are actually simplifying a lot our problem. The, we have seen that for artists and POIs, the connection is most of the times the result of a common node. This means that the direction must change at least one time. We reduced all the different cases to one. In fact what we do is to make two separate queries with depth  from the source and from the destination. These queries return the list of all half-paths. We make a set of the involved nodes and then we intersect the two sets, in order to find the common nodes. At this point we operate joins that create full paths. The whole matching of POIs and Artists,100.000 combinations, requires only some hours against the weeks that would take relfinder.
\subsection{Selection Algorithm}
Now that we have all our path, we need to find the best ones, those that shows interesting links between POIs and Artists.
todo: lorenzo
